$Z^\prime$ mass limits and the naturalness of supersymmetry

### 2015-05-24 paper/e6ssm-tuning-paper.tex
changed the way we describe the bilinear mass term.

The full superpotential for $E_6$ inspired models will then be,
\be
W_{E_6} = W_0 + W_1 + W_2 , \label{Eq:e6-superpot}
\ee
where
\begin{eqnarray}
W_0 &=& \lambda_{ijk} \hat{S}_i \hat{H}^d_{j} \cdot\hat{H}^u_{k} +
\kappa_{ijk} \hat{S}_i \hat{D}_j \hat{\overline{D}}_k + h^N_{ijk} \hat{N}^c_i
\hat{H}^u_{j} \cdot \hat{L}_k \nonumber \\
& & + y^U_{ijk} \hat{u}^c_i \hat{H}^u_{j} \cdot \hat{Q}_k + y^D_{ijk}
\hat{d}^c_i \hat{Q}_k \cdot \hat{H}^d_{j} + y^E_{ijk} \hat{e}^c_i \hat{L}_k
\cdot\hat{H}^d_{j} , \\
W_1 &=& g^Q_{ijk} \hat{D}_i \hat{Q}_j \cdot \hat{Q}_k + g^q_{ijk}
\hat{\bar{D}}_i \hat{d}^c_j \hat{u}^c_k , \\
W_2 &=& g^N_{ijk} \hat{N}^c_i \hat{D}_j \hat{d}^c_k + g^E_{ijk} \hat{e}^c_i
\hat{D}_j \hat{u}^c_k + g^D_{ijk} \hat{Q}_i \cdot \hat{L}_j \hat{\bar{D}}_k .
\label{Eq:e6-superpot-parts}
\end{eqnarray}
Nonetheless while this model is very elegant so far, the
superpotential of Eq.~(\ref{Eq:e6-superpot}) contains dangerous terms which
can induce proton decay and lead to large flavour changing neutral
currents (FCNCs).  There are a number of approaches to suppress these
terms, involving the use of different discrete symmetries.  Here for
the purposes of renormalization group running we will simply include
the following unsuppressed superpotential terms, which follows the
approach taken in work on the cE$_6$SSM \cite{Athron:2009ue, Athron:2009bs},
%
%% \begin{align}
%%   \begin{split}
    \ba
    W &\approx& y_{\tau} \SuperField{L}_3 \cdot \SuperField{H}_d
      \SuperField{e}^c_3 + y_b \SuperField{Q}_3\cdot\SuperField{H}_d
      \SuperField{d}_3^c + y_t \SuperField{H}_u\cdot\SuperField{Q}_3
      \SuperField{u}_3^c\nonumber\\
    &&
    + \lambda_i \SuperField{S} \SuperField{H}_i^d
    \cdot \SuperField{H}_i^u  + \kappa_i \SuperField{S} \SuperField{D}_i
    \SuperField{\overline{D}}_i + \mu' \SuperField{H}'\cdot
    \SuperField{\overline{H'}},
    \ea
  %% \end{split}
  \label{SuPot_RGE}
%%\end{align}
%
where we denote by $\SuperField{H}_d^3 \equiv \SuperField{H}_d$,
$\SuperField{H}_u^3 \equiv \SuperField{H}_u$ and $\SuperField{S}_3 \equiv
\SuperField{S}$ the third generation Higgs and SM singlet fields that are
assumed to acquire non-zero VEVs.  In addition to the $E_6$ preserving terms
following from Eq.~(\ref{Eq:e6-superpot}), this superpotential also contains a
bilinear term $\mu' \SuperField{H}'\cdot \SuperField{\overline{H'}}$ that
violates the $E_6$ symmetry but is invariant with respect to the low energy SM
gauge group and the additional $U(1)'$ symmetry.  This term is responsible for
setting the masses of the components of the superfields $\SuperField{H}'$,
$\SuperField{\overline{H'}}$, but it is not involved in the process of EWSB.
Consequently the impact on the fine tuning of the value of $\mu'$ is much
smaller than that coming from other sectors, and so can be safely neglected
in our study. In all of the scans we present below the value of $\mu'$ is
fixed to $\mu' = 5$ TeV.

@@

In this case we now consider realistic scenarios, where the
parameters that are not scanned over are set to values which keep the
associated states comfortably above their experimental limits.  So in
both the MSSM and E$_6$SSM, all other soft scalar masses are set
to $5$ TeV.  We also impose several theoretical constraints on the points
included in the analysis.  In particular, we require a valid spectrum with no
tachyonic states to exclude points which would have an unrealistic minimum,
for example due to the appearance of charge or color breaking (CCB) minima.
We work in the third family approximation, taking the first and second
generation Yukawa couplings to be zero, and we also assume their associated
soft trilinears vanish.  Similarly we take $A_b = A_\tau = 0$ GeV.  The $U(1)$
gaugino soft mass $M_1$ was fixed to $M_1 = 300$ GeV, and we fix $M_3 = 2000$
GeV.  Additionally, in the E$_6$SSM the $U(1)_N$ gaugino soft mass $M_1'$ is
held fixed with $M_1' = M_1 = 300$ GeV, and $\mu^\prime = 5$ TeV.

@@

There is another limit which plays a similar role.  Chargino limits
directly constrain the $\mu$ parameter (or effective $\mu$ parameter
in these $U(1)$ extensions).  The LEP bound \cite{Kraan:2005vy} on
chargino masses, excluding $m_{\tilde{\chi}^\pm_1} \lesssim 104$ GeV, implies
that $|\mu|$ should only be greater than $\sim 100$ GeV, which is not
substantially larger than $M_Z$.  Consequently the bound from LEP is not
high enough to have an impact on the fine tuning obtained in the models and
parameter space regions that we have studied, as we have checked explicitly.
Significantly larger lower bounds on the $\mu$ parameter, and therefore on the
fine tuning, may arise from chargino limits coming from LHC searches.  However
the chargino limits from the LHC depend on whether there are light sleptons or
sneutrinos and the mass difference between the lightest chargino and lightest
neutralino.  Current limits placed by CMS and ATLAS extend up to
$m_{\tilde{\chi}^\pm_1} \approx 700 - 740$ GeV if there are light
sleptons \cite{Khachatryan:2014qwa, Aad:2014nua} with much weaker
bounds if there are no light sleptons or sneutrinos\footnote{Useful
summary plots of these limit may be found on the public pages of
ATLAS, \url{https://atlas.web.cern.ch/Atlas/GROUPS/PHYSICS/CombinedSummaryPlots/SUSY/ATLAS_SUSY_EWSummary/ATLAS_SUSY_EWSummary.png}
and CMS \url{http://cms.web.cern.ch/sites/cms.web.cern.ch/files/styles/large/public/field/image/Image_03_exclusion_Combined.png?itok=8FMBpu_1}.}.

>>>>

The full superpotential for $E_6$ inspired models coming from $27\otimes 27 \otimes 27$ decomposition of the fundamental $E_6$ representation will then be,
\be
W_{E_6} = W_0 + W_1 + W_2 , \label{Eq:e6-superpot}
\ee
where
\begin{eqnarray}
W_0 &=& \lambda_{ijk} \hat{S}_i \hat{H}^d_{j} \cdot\hat{H}^u_{k} +
\kappa_{ijk} \hat{S}_i \hat{D}_j \hat{\overline{D}}_k + h^N_{ijk} \hat{N}^c_i
\hat{H}^u_{j} \cdot \hat{L}_k \nonumber \\
& & + y^U_{ijk} \hat{u}^c_i \hat{H}^u_{j} \cdot \hat{Q}_k + y^D_{ijk}
\hat{d}^c_i \hat{Q}_k \cdot \hat{H}^d_{j} + y^E_{ijk} \hat{e}^c_i \hat{L}_k
\cdot\hat{H}^d_{j} , \\
W_1 &=& g^Q_{ijk} \hat{D}_i \hat{Q}_j \cdot \hat{Q}_k + g^q_{ijk}
\hat{\bar{D}}_i \hat{d}^c_j \hat{u}^c_k , \\
W_2 &=& g^N_{ijk} \hat{N}^c_i \hat{D}_j \hat{d}^c_k + g^E_{ijk} \hat{e}^c_i
\hat{D}_j \hat{u}^c_k + g^D_{ijk} \hat{Q}_i \cdot \hat{L}_j \hat{\bar{D}}_k .
\label{Eq:e6-superpot-parts}
\end{eqnarray}
Nonetheless while this model is very elegant so far, the
superpotential of Eq.~(\ref{Eq:e6-superpot}) contains dangerous terms which
can induce proton decay and lead to large flavour changing neutral
currents (FCNCs).  There are a number of approaches to suppress these
terms, involving the use of different discrete symmetries.  Here for
the purposes of renormalization group running we will simply include
the following unsuppressed superpotential terms, which follows the
approach taken in work on the cE$_6$SSM \cite{Athron:2009ue, Athron:2009bs},
%
%% \begin{align}
%%   \begin{split}
    \ba
    W &\approx& y_{\tau} \SuperField{L}_3 \cdot \SuperField{H}_d
      \SuperField{e}^c_3 + y_b \SuperField{Q}_3\cdot\SuperField{H}_d
      \SuperField{d}_3^c + y_t \SuperField{H}_u\cdot\SuperField{Q}_3
      \SuperField{u}_3^c\nonumber\\
    &&
    + \lambda_i \SuperField{S} \SuperField{H}_i^d
    \cdot \SuperField{H}_i^u  + \kappa_i \SuperField{S} \SuperField{D}_i
    \SuperField{\overline{D}}_i + \mu' \SuperField{H}'\cdot
    \SuperField{\overline{H'}},
    \ea
  %% \end{split}
  \label{SuPot_RGE}
%%\end{align}
%
where we denote by $\SuperField{H}_d^3 \equiv \SuperField{H}_d$,
$\SuperField{H}_u^3 \equiv \SuperField{H}_u$ and $\SuperField{S}_3
\equiv \SuperField{S}$ the third generation Higgs and SM singlet
fields that are assumed to acquire non-zero VEVs.  In addition to the
terms coming from the $27\otimes 27 \otimes 27$ interactions given in
Eq.~(\ref{Eq:e6-superpot}), this superpotential also contains a
bilinear term $\mu' \SuperField{H}'\cdot \SuperField{\overline{H'}}$,
arising from $27^\prime \otimes \overline{27}^\prime$, which is
invariant with respect to the low energy SM gauge group and the
additional $U(1)'$ symmetry and also anomaly free.  This term is
responsible for setting the masses of the components of the
superfields $\SuperField{H}'$, $\SuperField{\overline{H'}}$, included
to ensure gauge coupling unification, but it is not involved in the
process of EWSB.  Consequently the impact on the fine tuning of the
value of $\mu'$ is much smaller than that coming from other sectors,
and so can be safely neglected in our study. In all of the scans we
present below the value of $\mu'$ is fixed to $\mu' = 5$ TeV.

@@

In this case we now consider realistic scenarios, where the
parameters that are not scanned over are set to values which keep the
associated states comfortably above their experimental limits.  So in
both the MSSM and E$_6$SSM, all other soft scalar masses are set
to $5$ TeV.  We also impose several theoretical constraints on the points
included in the analysis.  In particular, we require a valid spectrum with no
tachyonic states to exclude points which would have an unrealistic minimum,
for example due to the appearance of charge or color breaking (CCB) minima.
As a further requirement in the MSSM we exclude any points that do not satisfy
the well known CCB constraint \cite{Casas:1995pd},
\begin{equation} \label{eq:MSSMCCBconstraint}
A_t^2 < 3 \left ( \mu^2 + m_{H_u}^2 + m_{Q_3}^2 + m_{u_3}^2 \right ),
\end{equation}
or the unbounded from below (UFB) constraint,
\begin{equation} \label{eq:MSSMUFBconstraint}
m_{H_d}^2 + m_{H_u}^2 + 2 \mu^2 > 2 B \mu .
\end{equation}
We work in the third family approximation, taking the first and second
generation Yukawa couplings to be zero, and we also assume their associated
soft trilinears vanish.  Similarly we take $A_b = A_\tau = 0$ GeV.  The $U(1)$
gaugino soft mass $M_1$ was fixed to $M_1 = 300$ GeV, and we fix $M_3 = 2000$
GeV.  Additionally, in the E$_6$SSM the $U(1)_N$ gaugino soft mass $M_1'$ is
held fixed with $M_1' = M_1 = 300$ GeV, and $\mu^\prime = 5$ TeV.

@@

There is another limit which plays a similar role.  Chargino limits
directly constrain the $\mu$ parameter (or effective $\mu$ parameter
in these $U(1)$ extensions).  The LEP bound \cite{Kraan:2005vy} on
chargino masses, excluding $m_{\tilde{\chi}^\pm_1} \lesssim 104$ GeV, implies
that $|\mu|$ should only be greater than $\sim 100$ GeV, which is not
substantially larger than $M_Z$.  Consequently the bound from LEP is not
high enough to have an impact on the fine tuning obtained in the models and
parameter space regions that we have studied, as we have checked explicitly\footnote{It is for this reason that we did not require the LEP bounds be
satisfied for the points shown in Figure \ref{Fig:e6ssmvsmssm}
and in Figure \ref{Fig:othere6modelsvsmssm} below.}.  Significantly larger
lower bounds on the $\mu$ parameter, and therefore on the fine tuning, may
arise from chargino limits coming from LHC searches.  However the chargino
limits from the LHC depend on whether there are light sleptons or sneutrinos
and the mass difference between the lightest chargino and lightest neutralino.
Current limits placed by CMS and ATLAS extend up to
$m_{\tilde{\chi}^\pm_1} \approx 700 - 740$ GeV if there are light
sleptons \cite{Khachatryan:2014qwa, Aad:2014nua} with much weaker
bounds if there are no light sleptons or sneutrinos\footnote{Useful
summary plots of these limit may be found on the public pages of
ATLAS, \url{https://atlas.web.cern.ch/Atlas/GROUPS/PHYSICS/CombinedSummaryPlots/SUSY/ATLAS_SUSY_EWSummary/ATLAS_SUSY_EWSummary.png}
and CMS \url{http://cms.web.cern.ch/sites/cms.web.cern.ch/files/styles/large/public/field/image/Image_03_exclusion_Combined.png?itok=8FMBpu_1}.}.
